Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-04-01 SupportingInformation.Rmd
adding one line description before each r chunk stating what it does

<!--Basic setup-->
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
seed<-101
# families<-c(1,3:10,13,14,16:20)
source("mtime.R") #A function needed for caching
library(captioner)
```

```{r cop_fig_intro,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
library(copula)
library(VineCopula)
library(mvtnorm)
set.seed(seed)

source("copsurrognd.R")

#x <- read.table("./Data/dataset/x.txt")
#y <- read.table("./Data/dataset/y.txt")
#mat<-cbind(x,y)
#colnames(mat)<-c("x","y")

#***constants
N<-200 #length of time series
n<-2 #number of sampling locations
numdat<-1 #the number of datasets to get from the normal copula
rho<-0.8 #normal-copula parameter for dependence between locations
#this is a covariance, see below call to rmvnorm

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
gshape<-7.5
gscale<-1 #gamma parameters

cop_to_marg_trans<-function(x){
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
  #should be something skewed
}
# pnorm(m) makes normal to uniform
# then qgamma(...) makes uniform to gamma distributed
m<-cop_to_marg_trans(pnorm(m))

ccop<-claytonCopula(2,2)
csurrog<-array(NA,dim(m))
csurrog[,,1]<-copsurrognd(m[,,1],targetcop=ccop,numsurrog=1)
mat<-csurrog[,,1]

pdf("./Results/rawdata.pdf",height = 4, width = 4)
scale<-2
xhist <- hist(mat[,1], breaks=30, plot=FALSE)
yhist <- hist(mat[,2], breaks=30, plot=FALSE) 
top <- max(c(xhist$counts, yhist$counts)) 
xrange <- range(mat[,1])
yrange <- range(mat[,2])
nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0))  
plot(mat[,1], mat[,2], xlim=xrange, ylim=yrange, xlab="x", ylab="y", cex.lab=1.5)
legend("bottomright","raw data plot",bty="n",cex=1.5,text.col="magenta")
op<-par(mar=c(0,3,1,1)) 
barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
op<-par(mar=c(3,0,1,1)) 
barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
par(op)
dev.off()

pdf("./Results/copdata.pdf",height = 4, width = 4)
copmat<-pobs(mat)
xhist <- hist(copmat[,1], breaks=30, plot=FALSE)
yhist <- hist(copmat[,2], breaks=30, plot=FALSE) 
top <- max(c(xhist$counts, yhist$counts)) 
xrange <- c(0,1)
yrange <- c(0,1)
nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0)) 
plot(copmat[,1], copmat[,2], xlim=xrange, ylim=yrange, xlab="u=rank(x)/(n+1)", ylab="v=rank(y)/(n+1)",cex.lab=1.5)
legend("bottomright","copula plot",bty="n",cex=1.5,text.col="magenta")
op<-par(mar=c(0,3,1,1)) 
barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
op<-par(mar=c(3,0,1,1)) 
barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
par(op)
dev.off()
```

```{r cop_pedagog_fig,echo=F,results="hide"}
set.seed(seed)
source("./FirstPedagogFig.R")
source("./SecondPedagogFig.R")
source("./ThirdPedagogFig.R")
source("./MakeCopulaPedagogFigs.R")
```

<!--The raw data look like this:-->
```{r load_data_Loecke, echo=F, results='hide'}
#fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
d<-readRDS("Data/dataForShymolinaCoord.rds")
d<-d[,c("SOCstock100","TSNstock100","lat","long")]
d<-na.omit(d)
pdf("./Results/SoilCNraw.pdf")
op<-par(mar=c(6, 6, 0.2, 2.1))
plot(d[,1],d[,2],type="p",col="black",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=2.5)
par(op)
dev.off()
```

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, results='hide'}
source("getcopula.R")
#convert to a copula and plot
pdf("./Results/SoilCNcop.pdf")
v_CN<-getcopula(d=d[,c(1:2)],rankon=T,ploton=T) 
dev.off()
```

```{r plot_SoilCNmapUSA, echo=F, results='hide'}
#---------map of soilCN location on USA------------
library(maps)
pdf("./Results/SoilCNmapUSA.pdf",height=4,width=8)
map("usa")
points(x = d$long, y = d$lat, col="red",pch=16,cex=0.5)
dev.off()
```

```{r read_BMR_birds_and_copulaplot, echo=F, results='hide'}
library(readxl)
set.seed(seed)
raw_b<-readxl::read_xls("./Data/BMR/BirdBodyMassesMetabolicRates/BirdMassesAndMetabolicRatesMcNab2008.xls")
raw_b<-raw_b[,1:3]
raw_b<-na.omit(raw_b)
raw_b<-raw_b[,2:3]
raw_b<-as.data.frame(raw_b)

@@

```{r read_BMR_mammals_and_copulaplot, echo=F, results='hide'}
set.seed(seed)
raw_m<-readxl::read_xls("./Data/BMR/MammalianBodyMassesMetabolicRates/MammalianMassesAndMetabolicRatesMcNab2008.xls")
raw_m<-raw_m[,1:3]
raw_m<-na.omit(raw_m)
raw_m<-raw_m[,2:3]

@@

```{r read_ceder_creek_data, echo=F, results="hide"}
set.seed(seed)
#-------------------------------------------------
# This function takes 
# Input :
#       db : a dataframe : smaller subset of ple_120 data to calculate avg. biomass for each plot
#       dp : a dataframe : smaller subset of pce_120 data to calculate avg. shannon's H for those plots
#       year : the year for which you want to calculate biomass and shannon's index data
#       rawplot : logical : T for rawdata plot, F for copula plot
#       ploton : logical for optional plotting

@@

```{r map_plot_aphid_plankton,results="hide",echo=F}
source("aphid_data_calling.R")
source("good_loclist.R")
library(maps)

@@

```{r make_data_plot_fig,echo=FALSE,results="hide"}
source("./MakeBivarDataPlot.R")
```

@@

 [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had at least 45 years time series data (see Figure \ref{SM-fig_plankton_map}).

@@

```{r tab_aphid_info, echo=F, results='asis',message=F}
library(tinytex)
library(tibble)
library(kableExtra)
library(dplyr)
aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
                       "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
                       "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach potato aphid",
                       "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
                     c2=c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
                       "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",
                       "Brachycaudus helichrysi","Brevicoryne brassicae","Hyalopterus pruni","Acyrthosiphon pisum","Myzus persicae",
                       "Macrosiphum euphorbiae","Metopolophium dirhodum","Myzus ascalonicus","Drepanosiphum platanoidis","Cavariella aegopodii")
)

@@

```{r tab_plankton_info, echo=F, results='asis',message=F}
library(tibble)
plankton_info <- tibble(c1=c('Calanus stages I-IV','Para-pseudocalanus spp.','Temora longicornis','Acartia sp.',
                          'Centropages typicus',
                          'Oithona sp.','Echinoderm larvae','Calanus finmarchicus',
                          'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
                   'Euphausiids','Thalassiosira sp.','Rhizosolenia styliformis',
                   'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
                   'Proboscia alata','Pseudocalanus sp.',
                   'Pseudo-nitzschia delicatissima','Pseudo-nitzschia seriata')
)
cns<-c('Echinoderm larvae','Euphausiids','Decapoda larvae')
'%ni%' <- Negate('%in%')
idn<-which(plankton_info$c1%ni%cns)
knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 22 plankton taxa for which data were available \\label{tab_plankton_info}",
      col.names = c("Plankton taxon name"))%>%row_spec(row=idn,italic=T)
```
<!--***DAN: These names differ from those used in Emma's 2016 Glob Change Biol paper.
Need to check with Chris Reid. Have to get this right.-->

```{r read_methane_data, echo=F, results="markup"}
set.seed(seed)
methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
#class(methane_raw) #it's a data.frame for multivariable data
#knitr::kable(head(methane_raw,3),format='pandoc',
#             caption = "First few rows of raw methane data.\\label{tab_rawdata_methane}",
#             booktabs=T) # display first few rows
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/SoilCNmapUSA.pdf} 
\caption[Sampling locations for C and N data]{Locations of central pedons sampled 
for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
\end{center}
\end{figure}

@@

<!--The message=F and warning=F flags are to suppress warnings that come when you pass a parameter
of 0 to the Frank or Clayton copulas. These are technicallly boundary cases for these families.
They correspond to the independent copula. The packages we use provide the indepdent copula, but
issue a warning. Since anyway we want the independent copula, we can suppress the warning. We ran this
without the warning and it is OK to supress it.-->
<!-- The lines you are talking about appeared as message not as warning(actualy no warning appered for this chunk), so we have to set message=F here-->
```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
source("TestStats_multivar_dataset.R")
set.seed(seed)

@@

<!--Display model selection results in a table-->
```{r tab_soilCNfit, echo=F, results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h, #longtable = T, 
             format = "latex",linesep = "",
      caption.short = "Model fitting results for soil C and N dataset",
      caption = "Results of fitting our 16 copula families to the soil C and N dataset. Copula name abbreviations are given in the main text. AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
      booktabs = TRUE)
```

@@

<!--Now make and display a table, this one about asymmetry of tail dependence-->
```{r make_table_stat_fn, echo=F}

fractxt<-function(fracval,numsurrog){
  ptxt<-''
  if (fracval>.5*numsurrog){                        # ?DCR why 0.5 factor?
    ptxt<-paste(">",fracval,sep='')
  } 
  if (fracval<=.5*numsurrog){
    ptxt<-paste("<",numsurrog-fracval,sep='')
  }
  return(ptxt)
}

@@

```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
             format = "latex",linesep = "",
      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons. A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
```

```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_birds<-OurBiCopSelect(u1=cop_b[,1],u2=cop_b[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_birds,file="./Results/BMR_results/birds/BivMS_birds.RDS")
```

@@

```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_birds_ga<-bivfunctionplot(v=cop_b,resloc="./Results/BMR_results/birds/",nametag="birds_numbin_10",numbin = 10)
```

@@

```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_birds_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
             booktabs=T)
```

```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_mammals<-OurBiCopSelect(u1=cop_m[,1],u2=cop_m[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_mammals,file="./Results/BMR_results/mammals/BivMS_mammals.RDS")
```

@@

```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_mammals_ga<-bivfunctionplot(v=cop_m,resloc="./Results/BMR_results/mammals/",nametag="mammals_numbin_10",numbin = 10)
```

@@

```{r tab_bmr_mammals_asym, echo=F, results="markup"}
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_mammals_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure. \\label{tab_bmr_mammals_asym}",
             booktabs=T)
```

```{r bivmodselect_cc, echo=F, cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)

@@

<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the Cedar Creek dataset",
             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
             booktabs=T)
```

```{r stat_cc_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)

>>>>

<!--Basic setup-->
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
seed<-101
source("mtime.R") #A function needed for caching
library(captioner)
```

<!--generating plots for pedagog figures-->
```{r cop_pedagog_fig,echo=F,results="hide"}
set.seed(seed)
source("./FirstPedagogFig.R")
source("./SecondPedagogFig.R")
source("./ThirdPedagogFig.R")
source("./MakeCopulaPedagogFigs.R")
```

<!--reading data and generaing plot for soilCN-->
```{r load_and_plot_data_soilCN, echo=F, results='hide'}
#fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
d<-readRDS("Data/dataForShymolinaCoord.rds")
d<-d[,c("SOCstock100","TSNstock100","lat","long")]
d<-na.omit(d)
pdf("./Results/SoilCNraw.pdf")
op<-par(mar=c(6, 6, 0.2, 2.1))
plot(d[,1],d[,2],type="p",col="black",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=2.5)
par(op)
dev.off()
```

<!--generating copula plot with soilCN data-->
```{r copula_plot_soilCN, echo=F, results='hide'}
source("getcopula.R")
#convert to a copula and plot
pdf("./Results/SoilCNcop.pdf")
v_CN<-getcopula(d=d[,c(1:2)],rankon=T,ploton=T) 
dev.off()
```

<!--generating plots with sampling locations for soilCN data-->
```{r plot_SoilCNmapUSA, echo=F, results='hide'}
#---------map of soilCN location on USA------------
library(maps)
pdf("./Results/SoilCNmapUSA.pdf",height=4,width=8)
map("usa")
points(x = d$long, y = d$lat, col="red",pch=16,cex=0.5)
dev.off()
```

<!--reading Birds' BMR data and generating raw, copula plots with it-->
```{r read_BMR_birds_and_copulaplot, echo=F, results='hide'}
library(readxl)
set.seed(seed)
raw_b<-readxl::read_xls("./Data/BMR/BirdBodyMassesMetabolicRates/BirdMassesAndMetabolicRatesMcNab2008.xls")
raw_b<-raw_b[,1:3]
raw_b<-na.omit(raw_b)
raw_b<-raw_b[,2:3]
raw_b<-as.data.frame(raw_b)

@@

<!--reading Mammals' BMR data and generating raw, copula plots with it-->
```{r read_BMR_mammals_and_copulaplot, echo=F, results='hide'}
set.seed(seed)
raw_m<-readxl::read_xls("./Data/BMR/MammalianBodyMassesMetabolicRates/MammalianMassesAndMetabolicRatesMcNab2008.xls")
raw_m<-raw_m[,1:3]
raw_m<-na.omit(raw_m)
raw_m<-raw_m[,2:3]

@@

<!--reading ceder creek data and generating raw and copula plots with it-->
```{r read_ceder_creek_data, echo=F, results="hide"}
set.seed(seed)
#-------------------------------------------------
# This function takes 
# Input :
#       db : a dataframe : smaller subset of ple_120 data to calculate avg. biomass for each plot
#       dp : a dataframe : smaller subset of pce_120 data to calculate avg. shannon's H for those plots
#       year : the year for which you want to calculate biomass and shannon's index data
#       rawplot : logical : T for rawdata plot, F for copula plot
#       ploton : logical for optional plotting

@@

<!--reading aphid and plankton data, then generating plots showing sampling locations-->
```{r map_plot_aphid_plankton,results="hide",echo=F}
source("aphid_data_calling.R")
source("good_loclist.R")
library(maps)

@@

<!--Generating a single plot containing raw and copula plots for all bivariate data-->
```{r make_data_plot_fig,echo=FALSE,results="hide"}
source("./MakeBivarDataPlot.R")
```

@@

 [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had at least 45 years time series data (see Figure \ref{fig_plankton_map}).

@@

<!--generating table with information about 20 aphid species-->
```{r tab_aphid_info, echo=F, results='asis',message=F}
library(tinytex)
library(tibble)
library(kableExtra)
library(dplyr)
aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
                       "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
                       "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach potato aphid",
                       "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
                     c2=c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
                       "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",
                       "Brachycaudus helichrysi","Brevicoryne brassicae","Hyalopterus pruni","Acyrthosiphon pisum","Myzus persicae",
                       "Macrosiphum euphorbiae","Metopolophium dirhodum","Myzus ascalonicus","Drepanosiphum platanoidis","Cavariella aegopodii")
)

@@

<!--generating table with information about 22 plankton species-->
```{r tab_plankton_info, echo=F, results='asis',message=F}
library(tibble)
plankton_info <- tibble(c1=c('Calanus stages I-IV','Para-pseudocalanus spp.','Temora longicornis','Acartia sp.',
                          'Centropages typicus',
                          'Oithona sp.','Echinoderm larvae','Calanus finmarchicus',
                          'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
                   'Euphausiids','Thalassiosira sp.','Rhizosolenia styliformis',
                   'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
                   'Proboscia alata','Pseudocalanus sp.',
                   'Pseudo-nitzschia delicatissima','Pseudo-nitzschia seriata')
)
cns<-c('Echinoderm larvae','Euphausiids','Decapoda larvae')
'%ni%' <- Negate('%in%')
idn<-which(plankton_info$c1%ni%cns)
knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 22 plankton taxa for which data were available \\label{tab_plankton_info}",
      col.names = c("Plankton taxon name"))%>%row_spec(row=idn,italic=T)
```

<!--reading multivariate methane flux data-->
```{r read_methane_data, echo=F, results="markup"}
set.seed(seed)
methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
```

<!--embedding plot with sampling locations : soilCN, aphids, planktons-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/SoilCNmapUSA.pdf} 
\caption[Sampling locations for C and N data]{Locations of central pedons sampled 
for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
\end{center}
\end{figure}

@@

<!--Simulation for testing non-parametric statistics with different copulas :
The message=F is to suppress warnings that come when you pass a parameter
of 0 to the Frank or Clayton copulas. These are technicallly boundary cases for these families.
They correspond to the independent copula. The packages we use provide the indepdent copula, but
issue a message.-->
```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
source("TestStats_multivar_dataset.R")
set.seed(seed)

@@

<!--Display model selection results for soilCN data in a table-->
```{r tab_soilCNfit, echo=F, results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h, #longtable = T, 
             format = "latex",linesep = "",
      caption.short = "Model fitting results for soil C and N dataset",
      caption = "Results of fitting our 16 copula families to the soil C and N dataset. Copula name abbreviations are given in the main text. AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
      booktabs = TRUE)
```

@@

<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
```{r make_table_stat_fn, echo=F}

fractxt<-function(fracval,numsurrog){
  ptxt<-''
  if (fracval>.5*numsurrog){                        
    ptxt<-paste(">",fracval,sep='')
  } 
  if (fracval<=.5*numsurrog){
    ptxt<-paste("<",numsurrog-fracval,sep='')
  }
  return(ptxt)
}

@@

<!--Displaying a table with npa summary results for soilCN data-->
```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
             format = "latex",linesep = "",
      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons. A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
```

<!--Simulations for model selection results with Birds' BMR data-->
```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_birds<-OurBiCopSelect(u1=cop_b[,1],u2=cop_b[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_birds,file="./Results/BMR_results/birds/BivMS_birds.RDS")
```

@@

<!--This chunk is for npa testing results for Birds' BMR dataset-->
```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_birds_ga<-bivfunctionplot(v=cop_b,resloc="./Results/BMR_results/birds/",nametag="birds_numbin_10",numbin = 10)
```

@@

<!--Displaying a table with npa summary results for Birds' BMR data-->
```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_birds_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
             booktabs=T)
```

<!--Simulations for model selection results with Mammals' BMR data-->
```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_mammals<-OurBiCopSelect(u1=cop_m[,1],u2=cop_m[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_mammals,file="./Results/BMR_results/mammals/BivMS_mammals.RDS")
```

@@

<!--This chunk is for npa testing results for Mammals' BMR dataset-->
```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_mammals_ga<-bivfunctionplot(v=cop_m,resloc="./Results/BMR_results/mammals/",nametag="mammals_numbin_10",numbin = 10)
```

@@

<!--Displaying a table with npa summary results for Mammals' BMR data-->
```{r tab_bmr_mammals_asym, echo=F, results="markup"}
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_mammals_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure. \\label{tab_bmr_mammals_asym}",
             booktabs=T)
```

<!--Simulations for model selection results with Cedar creek data-->
```{r bivmodselect_cc, echo=F, cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)

@@

<!--Display model selection results for Cedar-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the Cedar Creek dataset",
             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
             booktabs=T)
```
